---
title: "Exp. 1 - Measurement invariance"
toc: false
knitr:
  opts_chunk:
    fig.path: "figures/"
---

Measurement invariance across the vignettes was examined using multigroup CFA in lavaan with robust maximum likelihood estimation (MLR) and participant ID as a clustering variable. Configural, metric, and scalar invariance models were compared sequentially. Table 1 below summarises the invariance results, but generally model fit did not degrade substantially when factor loadings were constrained (ΔCFI = .001, ΔRMSEA = .008), suggesting metric invariance for FM, CB and AIA. However, when intercepts were constrained (ΔCFI < .005, ΔRMSEA < .004), all constructs failed to support the scalar invariance across vignettes.

These analysyses correspond to section `CFA model fits and measurement invariance` in the manuscript.


```{r}
#| echo: false
#| warning: false
#| message: false
#| id: libraries
library(tidyr)
library(dplyr)
library(stringr)
#library(here)
library(psych)
library(lavaan)
set.seed(3110)
```

## EDR

```{r}
#| echo: true
#| id: sem-edr

df <- readRDS(file="data/EDR_N100.rds")
tmp <- dplyr::filter(df,vignette=="E" | vignette=="D" | vignette=="R")
tmp_w <- pivot_wider(tmp, names_from = item_label, values_from = value,id_expand = FALSE,id_cols = c('ID','vignette'))
tmp_w<-data.frame(tmp_w)
rownames(tmp_w) <- 1:nrow(tmp_w)

model <- "
E =~ E1 + E6 + E7
D =~ D1 + D6 + D7
R =~ R1 + R3 + R7
"
# baseline
model.baseline <- lavaan::cfa(model, data = tmp_w,meanstructure = TRUE,estimator = "MLR")

# configural invariance
model.configural <- lavaan::cfa(model, data = tmp_w, group = "vignette",meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.baseline,model.configural)
#summary(s,fit.measures="all")
config1 <- lavTestLRT(model.baseline,model.configural)$`Pr(>Chisq)`[2]
# metric invariance
model.metric <- lavaan::cfa(model, data = tmp_w, group = "vignette",group.equal=c("loadings"),meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.configural, model.metric)
#summary(s,fit.measures="all")
metric1 <- lavTestLRT(model.configural,model.metric)$`Pr(>Chisq)`[2]

# scalar invariance
model.scalar <- lavaan::cfa(model, data = tmp_w, group = "vignette", group.equal=c("loadings","intercepts"),meanstructure = TRUE,estimator = "MLR")
#summary(model.scalar, standardized = TRUE, fit.measures = TRUE)
#s<-semTools::compareFit(model.metric, model.scalar)
#summary(s)
scalar1 <- lavTestLRT(model.metric,model.scalar)$`Pr(>Chisq)`[2]

s<-semTools::compareFit(model.baseline,model.configural, model.metric,model.scalar,nested = TRUE)
summary(s)

```

## FM

```{r}
#| echo: true
#| id: sem-fm
df <- readRDS(file="data/FM_N100.rds")

tmp <- dplyr::filter(df,vignette=="F" | vignette=="M")
tmp_w <- pivot_wider(tmp, names_from = item_label, values_from = value,id_expand = FALSE,id_cols = c('ID','vignette'))
tmp_w<-data.frame(tmp_w)
rownames(tmp_w) <- 1:nrow(tmp_w)
#tmp_w <- select(tmp_w, -ID)
tmp_w$ID<-as.numeric(as.factor(tmp_w$ID))
tmp_w$vignette<-as.factor(tmp_w$vignette)

model <- '
    M =~ M1 + M3 + F1
    F =~ F3 + F6 + F10
'

# baseline
model.baseline <- lavaan::cfa(model, data = tmp_w,meanstructure = TRUE,estimator = "MLR")

# configural invariance
model.configural <- lavaan::cfa(model, data = tmp_w, group = "vignette",meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.baseline,model.configural)
#summary(s,fit.measures="all")
config2 <- lavTestLRT(model.baseline,model.configural)$`Pr(>Chisq)`[2]
# metric invariance
model.metric <- lavaan::cfa(model, data = tmp_w, group = "vignette",group.equal=c("loadings"),meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.configural, model.metric)
#summary(s,fit.measures="all")
metric2 <- lavTestLRT(model.configural,model.metric)$`Pr(>Chisq)`[2]

# scalar invariance
model.scalar <- lavaan::cfa(model, data = tmp_w, group = "vignette", group.equal=c("loadings","intercepts"),meanstructure = TRUE,estimator = "MLR")
#summary(model.scalar, standardized = TRUE, fit.measures = TRUE)
#s<-semTools::compareFit(model.metric, model.scalar)
#summary(s)
scalar2 <- lavTestLRT(model.metric,model.scalar)$`Pr(>Chisq)`[2]

s<-semTools::compareFit(model.baseline,model.configural, model.metric,model.scalar,nested = TRUE)
summary(s)

```

## CB

```{r}
#| echo: true
#| id: sem-cb

df <- readRDS(file="data/CB_N100.rds")

tmp <- dplyr::filter(df,vignette=="C" | vignette=="S")
tmp_w <- pivot_wider(tmp, names_from = item_label, values_from = value,id_expand = FALSE,id_cols = c('ID','vignette'))
tmp_w<-data.frame(tmp_w)
rownames(tmp_w) <- 1:nrow(tmp_w)
tmp_w$ID<-as.numeric(as.factor(tmp_w$ID))
tmp_w$vignette<-as.factor(tmp_w$vignette)

model <- "
    C =~ C5 + C2 + C9
    S =~ S5 + S6 + S7 
"

# baseline
model.baseline <- lavaan::cfa(model, data = tmp_w,meanstructure = TRUE,estimator = "MLR")

# configural invariance
model.configural <- lavaan::cfa(model, data = tmp_w, group = "vignette",meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.baseline,model.configural)
#summary(s,fit.measures="all")
config3 <- lavTestLRT(model.baseline,model.configural)$`Pr(>Chisq)`[2]
# metric invariance
model.metric <- lavaan::cfa(model, data = tmp_w, group = "vignette",group.equal=c("loadings"),meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.configural, model.metric)
#summary(s,fit.measures="all")
metric3 <- lavTestLRT(model.configural,model.metric)$`Pr(>Chisq)`[2]

# scalar invariance
model.scalar <- lavaan::cfa(model, data = tmp_w, group = "vignette", group.equal=c("loadings","intercepts"),meanstructure = TRUE,estimator = "MLR")
#summary(model.scalar, standardized = TRUE, fit.measures = TRUE)
#s<-semTools::compareFit(model.metric, model.scalar)
#summary(s)
scalar3 <- lavTestLRT(model.metric,model.scalar)$`Pr(>Chisq)`[2]

s<-semTools::compareFit(model.baseline,model.configural, model.metric,model.scalar,nested = TRUE)
summary(s)

```

## PEP

```{r}
#| echo: true
#| id: sem-pep

df <- readRDS(file="data/PEP_N100.rds")

tmp <- dplyr::filter(df,vignette=="U" | vignette=="P")
tmp_w <- pivot_wider(tmp, names_from = item_label, values_from = value,id_expand = FALSE,id_cols = c('ID','vignette'))
tmp_w<-data.frame(tmp_w)
rownames(tmp_w) <- 1:nrow(tmp_w)
#tmp_w <- select(tmp_w, -ID)
tmp_w$ID<-as.numeric(as.factor(tmp_w$ID))
tmp_w$vignette<-as.factor(tmp_w$vignette)

model <- "
    U =~ U2 + U3 + U6
    P =~ P4 + P8 + P9
"


# baseline
model.baseline <- lavaan::cfa(model, data = tmp_w,meanstructure = TRUE,estimator = "MLR")

# configural invariance
model.configural <- lavaan::cfa(model, data = tmp_w, group = "vignette",meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.baseline,model.configural)
#summary(s,fit.measures="all")
config4 <- lavTestLRT(model.baseline,model.configural)$`Pr(>Chisq)`[2]
# metric invariance
model.metric <- lavaan::cfa(model, data = tmp_w, group = "vignette",group.equal=c("loadings"),meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.configural, model.metric)
#summary(s,fit.measures="all")
metric4 <- lavTestLRT(model.configural,model.metric)$`Pr(>Chisq)`[2]

# scalar invariance
model.scalar <- lavaan::cfa(model, data = tmp_w, group = "vignette", group.equal=c("loadings","intercepts"),meanstructure = TRUE,estimator = "MLR")
#summary(model.scalar, standardized = TRUE, fit.measures = TRUE)
#s<-semTools::compareFit(model.metric, model.scalar)
#summary(s)
scalar4 <- lavTestLRT(model.metric,model.scalar)$`Pr(>Chisq)`[2]

s<-semTools::compareFit(model.baseline,model.configural, model.metric,model.scalar,nested = TRUE)
summary(s)

```

## AIA

```{r}
#| echo: true
#| id: sem-aia

df <- readRDS(file="data/AIA_N101.rds")

tmp <- dplyr::filter(df,vignette=="I" | vignette=="X" | vignette=="B")
tmp_w <- pivot_wider(tmp, names_from = item_label, values_from = value,id_expand = FALSE,id_cols = c('ID','vignette'))
tmp_w<-data.frame(tmp_w)
rownames(tmp_w) <- 1:nrow(tmp_w)
tmp_w$ID<-as.numeric(as.factor(tmp_w$ID))
tmp_w$vignette<-as.factor(tmp_w$vignette)

model <- "
    I =~ I3 + I4 + I5
    X =~ X1 + X2
    B =~ B1 + B3 + B5
"


# baseline
model.baseline <- lavaan::cfa(model, data = tmp_w,meanstructure = TRUE,estimator = "MLR")

# configural invariance
model.configural <- lavaan::cfa(model, data = tmp_w, group = "vignette",meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.baseline,model.configural)
#summary(s,fit.measures="all")
config5 <- lavTestLRT(model.baseline,model.configural)$`Pr(>Chisq)`[2]
# metric invariance
model.metric <- lavaan::cfa(model, data = tmp_w, group = "vignette",group.equal=c("loadings"),meanstructure = TRUE,estimator = "MLR")
#s<-semTools::compareFit(model.configural, model.metric)
#summary(s,fit.measures="all")
metric5 <- lavTestLRT(model.configural,model.metric)$`Pr(>Chisq)`[2]

# scalar invariance
model.scalar <- lavaan::cfa(model, data = tmp_w, group = "vignette", group.equal=c("loadings","intercepts"),meanstructure = TRUE,estimator = "MLR")
#summary(model.scalar, standardized = TRUE, fit.measures = TRUE)
#s<-semTools::compareFit(model.metric, model.scalar)
#summary(s)
scalar5 <- lavTestLRT(model.metric,model.scalar)$`Pr(>Chisq)`[2]

s<-semTools::compareFit(model.baseline,model.configural, model.metric,model.scalar,nested = TRUE)
summary(s)

```

## Summary

```{r}
#| echo: false
#| eval: true
#| id: compile-results
#| output: asis
options(scipen=999)
# Compile a table of results from all analyses above
results_table1 <- rbind(
  configural = c(config1,config2,config3,config4,config5),
  metric = c(metric1,metric2,metric3,metric4,metric5),
  scalar = c(scalar1,scalar2,scalar3,scalar4,scalar5)
)
rownames(results_table1) <- c("Configural","Metric","Scalar")
colnames(results_table1) <- c("EDR","FM","CB","PEP","AIA")

print(knitr::kable(t(results_table1),digits=3,caption="Measurement Invariance p-values for robust CFA analysis across constructs."))


```



```{r}
#| eval: false
#| echo: false
#| id: item-invar

library(misty)
df <- readRDS(file="data/EDR_N100.rds")
tmp <- dplyr::filter(df,vignette=="E" | vignette=="D" | vignette=="R")
tmp_w <- pivot_wider(tmp, names_from = item_label, values_from = value,id_expand = FALSE,id_cols = c('ID','vignette'))
tmp_w<-data.frame(tmp_w)
rownames(tmp_w) <- 1:nrow(tmp_w)
#tmp_w <- select(tmp_w, -ID)
model.test1 = list(factor1 = c("E1", "E6", "E7"), factor2 = c("D1", "D6", "D7"), factor3 = c("R1", "R3", "R7"))
invariance1 <- item.invar(data=tmp_w,model=model.test1,
                          group="vignette",
                          estimator="MLR",
                          print.fit="robust")
#invariance2 <- item.invar(data=ratings,model=model.test2,group="vignette")
invariance1$model.fit$metric

```





