---
title: "Exp. 1 - Preprocessing"
toc: false
knitr:
  opts_chunk:
    fig.path: "figures/"
---


```{r}
#| echo: false
#| warning: false
#| messages: false
#| id: setup
library(tidyr)
library(dplyr)
library(stringr)
library(here)
library(psych)
library(ggplot2)

source('scr/convert_items_long.R') # custom function
source('scr/participant_reliability.R') # custom function
source('scr/CFA_averaged.R') # custom function
```

## Load and preprocess data

```{r}
#| echo: true
#| id: loadata
source('scr/read_EFA_qualtrics.R') # custom function
fn <- "data/EDR - Pilot_August 18, 2025_10.15.csv"
d1 <- read_EFA_qualtrics(filename = fn, subconstructs = c("E","D","R"))
df1<- convert_items_long(d1)
df1b<- participant_reliability(df1,THRESHOLD = 0.001)
df1b$study <- 1

fn <- "data/FM_August 27, 2025_04.04.csv"
d2 <- read_EFA_qualtrics(filename=fn,subconstructs = c("F","M"), controlQ = "M_15")
df2<- convert_items_long(d2)
df2b<- participant_reliability(df2,THRESHOLD = 0.001)
df2b$study <- 2

fn <- "data/CB - redo_September 1, 2025_03.36.csv"
d3 <- read_EFA_qualtrics(filename=fn,subconstructs = c("C","S"), controlQ = "C_16")
df3<- convert_items_long(d3)
df3b<- participant_reliability(df3,THRESHOLD = 0.001)
df3b$study <- 3

fn <- "data/PEP_September 2, 2025_07.17.csv"
d4 <- read_EFA_qualtrics(filename=fn,subconstructs = c("U","P"), controlQ = "U_21")
df4<- convert_items_long(d4)
df4b<- participant_reliability(df4,THRESHOLD = 0.001)
df4b$study <- 4

fn <- "data/AIA_September 3, 2025_05.02.csv"
d5 <- read_EFA_qualtrics(filename=fn,subconstructs = c("I","X","B"), controlQ = "I_17")
df5<- convert_items_long(d5)
df5b<- participant_reliability(df5,THRESHOLD = 0.001)
df5b$study <- 5

```

## Demographics

Setting up demographic analysis.

```{r}
demographics <- rbind(df1b, df2b, df3b, df4b, df5b)
demo_clean <- demographics[!duplicated(demographics$ID),]

#aov.age <- aov(demo_clean$study ~ demo_clean$Age, data = demo_clean) # this is wrong
aov.age <- aov(Age ~ factor(study), data = demo_clean) # this is correct
summary(aov.age)

demo_clean<-dplyr::filter(demo_clean,Gender!="Non-binary")
ch.gender <- chisq.test(demo_clean$Gender, factor(demo_clean$study),simulate.p.value = FALSE)
print(knitr::kable(table(demo_clean$Gender, demo_clean$study))) 
ch.gender

M <- as.table(rbind(c(69, 31), c(50, 50))) # extreme group
print(chisq.test(M))
M <- as.table(rbind(c(41, 58), c(50, 50)))
chisq.test(M)


table(demo_clean$OMSI)

ch.OMSI <- chisq.test(demo_clean$OMSI, factor(demo_clean$study),simulate.p.value = FALSE) # Correct
ch.OMSI
table(demo_clean$OMSI,demo_clean$study)

#aov.OMSI <- aov(demo_clean$study ~ demo_clean$OMSI, data = demo_clean)

#OMSI <- c("Nonmusician" = 1, "Music-loving Nonmusician" = 2, "Amateur musician" = 3, "Serious amateur musician" = 4, "Semiprofessional musician" = 5, "Professional musician" = 5)

#table(demo_clean$Gender)
#Gender <- c("Female" = 1, "Male" = 2, "Non-binary" = 3, "Other" = 4, "Prefer not to say" = 5)

```

## Looking at Relevance

```{r}

EDRrel <- c(d1$D_Rep, d1$R_Rel, d1$E_Rel)
FMrel <- c(d2$M_Rel, d2$F_Rel)
CBrel <- c(d3$S_Rel, d3$C_Rel)
PEPrel <- c(d4$P_Rel, d4$U_Rel)
AIArel <- c(d5$B_Rel, d5$X_Rel, d5$I_Rel)

```

## Save preprocessed data

Before running any structural analysis, let's revise the item names.

```{r}
#| echo: true
#| output: asis
#| id: save-data
save_requested <- FALSE
if(save_requested==TRUE){
  saveRDS(df1b,file="data/EDR_N100.rds")
  saveRDS(df2b,file="data/FM_N100.rds")
  saveRDS(df3b,file="data/CB_N100.rds")
  saveRDS(df4b,file="data/PEP_N100.rds")
  saveRDS(df5b,file="data/AIA_N101.rds")
}
```

# Looking at various reliability estimates

From Revelle & Condon, 2019; Shrout & Lane, 2012. See Vispoel, Morris, & Kilinc, 2018; ten Hove, Jorgensen, & van der Ark, 2022

```{r}
#| echo: true
#| eval: false
#| id: relability
source('scr/ml_r.R') #just multilevel.reliability but works, what is the good value?
#ml_r(df1b) #RkF = 0.97, R1R = 0.37, RkR = 0.64, Rc = 0.92, RkRn = 0.58, Rcn = 0.93
#ml_r(df2b) #RkF = 0.95, R1R = 0.40, RkR = 0.57, Rc = 0.91, RkRn = 0.52, Rcn = 0.84
#ml_r(df3b) #RkF = 0.96, R1R = 0.59, RkR = 0.75, Rc = 0.85, RkRn = 0.72, Rcn = 0.85
#ml_r(df4b) #RkF = 0.98, R1R = 0.63, RkR = 0.78, Rc = 0.88, RkRn = 0.76, Rcn = 0.89
#ml_r(df5b) #RkF = 0.99, R1R = 0.70, RkR = 0.87, Rc = 0.86, RkRn = 0.87, Rcn = 0.76

#These are intraclass correlations, I assume they can be interpreted as such;
# 0 - .5 = poor; .5 - .75 moderate; .75 - .9 = good; above .9 = excellent

#trying to specify in our language what this means ------------------------------------------------
#RkF is the reliability of average ratings across all items and times, similar to Alpha
#R1R is generalisability of single time point across all items, specific vignette reliability, reliability of a single time
#RKR is generalisability of average time points across all items, person mean across vignettes
#Rc is generalisability of change scores, within person variability?
#RkRn is generalisability of between person differences averaged over time (nested within people)
#Rcn is generalisability of within person differences, averaged over items: reliability between people over time
```

# Factor analyses

```{r}
#| echo: true
#| eval: false
#| id: CFA
source('scr/CFA_averaged.R') # custom function
FA1 <- CFA_averaged(data=df1b,vignettes = c("E","D","R"))

# fix reverse scored items in the FM first
df2b$value[df2b$item_label=="F8"] <- 6 - df2b$value[df2b$item_label=="F8"]

FA2 <- CFA_averaged(data=df2b,vignettes=c("F","M"))
FA3 <- CFA_averaged(data=df3b,vignettes=c("C","S"))
FA4 <- CFA_averaged(data=df4b,vignettes=c("U","P"))
FA5 <- CFA_averaged(data=df5b,vignettes=c("I","X","B"))

```

### Visualise the mean ratings

```{r}
#| echo: true
#| id: visualise
#| fig.height: 10
#| fig.widht: 12

source('scr/visualise_means.R')
fig1 <- visualise_means(df1b)
fig2 <- visualise_means(df2b)
fig3 <- visualise_means(df3b)
fig4 <- visualise_means(df4b)
fig5 <- visualise_means(df5b)

```
