---
title: "Exp. 2 - Confirmatory Factor Analyses"
toc: false
knitr:
  opts_chunk:
    fig.path: "figures/"
---

```{r}
#| echo: false
#| warning: false
#| messages: false
#| eval: true
#| id: exp2-setup
library(tidyr)
library(dplyr)
library(stringr)
library(here)
library(psych)
library(ggplot2)
```

## Load data

```{r}
#| echo: true
#| id: exp2-load-data-rds
df_f <- readRDS("data/items35_N1200.rds")
source('scr/rename_specific_items.R') # We decided to relabel some constructs for clarity
df_f <- rename_specific_items(df_f)

# F1 should be in Motivation, relabelled as M9
df_f$item_label[df_f$item_label=="F1"] <- "M9"


```

```{r}
#| echo: false
#| output: asis
#| id: exp2-vignette-labels

df_f$itemCategory<-NA
df_f$itemCategory[str_detect(df_f$item_label,"E|D|R")]<-"EDR"
df_f$itemCategory[str_detect(df_f$item_label,"M|F")]<-"FM"
df_f$itemCategory[str_detect(df_f$item_label,"G|L")]<-"CB"     # Labels updated
df_f$itemCategory[str_detect(df_f$item_label,"C|X")]<-"PEP"    # Labels updated
df_f$itemCategory[str_detect(df_f$item_label,"S|I|B")]<-"AIA"  # Labels updated

# Check the assignments
#table(df_f$item_label,df_f$itemCategory)

```

## Report CFA Model Fit for each Construct

For each construct, report (1) CFA fit, (2) measurement invariance fit indices (metric, scalar), and (3) discriminant validity indices using the nested model comparison, $\chi^2$ (merge), see Rönkkö & Cho (2022).

### CFA - EDR

```{r}
#| echo: true
#| id: exp2-cfa-edr
#| warning: false
#| messages: false

library(lavaan)
library(semTools)
set.seed(3110)
source('scr/CFA_items.R')

EDR <- "
E =~ E1 + E6 + E7
D =~ D1 + D6 + D7
R =~ R1 + R3 + R7
"

items <- c("E1","E6","E7","D1","D6","D7","R1","R3","R7")
tmp <- CFA_items(df_f,vignette = c(1,2,3), items=items,vignette_name="EDR",bonusquestion="HAAS")
fit <- lavaan::cfa(model = EDR, data = tmp, estimator = "MLR",group = "VigNro") # includes vignettes explicitly
fit.meas1 <- lavaan::fitMeasures(fit,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit.meas1)
fit_metric <- lavaan::cfa(model = EDR, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = "loadings") # includes 
lavaan::lavTestLRT(fit, fit_metric)
fit_scalar <- lavaan::cfa(model = EDR, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = c("loadings","intercepts")) # includes 
lavaan::lavTestLRT(fit, fit_scalar)


fit <- lavaan::cfa(model = EDR, data = tmp, meanstructure = TRUE, estimator = "MLR")  # ignores vignettes
ds1 <- semTools::discriminantValidity(object = fit,merge = TRUE)
print(ds1)

```

### CFA - FM

```{r}
#| echo: true
#| id: exp2-cfa-fm
set.seed(3110)

FM <- "
M =~ M1 + M3 + M9 # Note one Focus item (F1) has been renamed to motivation item (M9) earlier
F =~ F3 + F6 + F10
"

items <- c("M1","M3","M9","F3","F6","F10")

tmp <- CFA_items(df_f,vignette = c(4,5), items=items,vignette_name="FM",bonusquestion="HAAS")
fit <- lavaan::cfa(model = FM, data = tmp, estimator = "MLR",group = "VigNro") # includes vignettes explicitly
fit.meas2 <- lavaan::fitMeasures(fit,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit.meas2)

fit_metric <- lavaan::cfa(model = FM, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = "loadings") # includes 
lavaan::lavTestLRT(fit, fit_metric)
fit_scalar <- lavaan::cfa(model = FM, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = c("loadings","intercepts")) # includes 
lavaan::lavTestLRT(fit, fit_scalar)


fit <- lavaan::cfa(model = FM, data = tmp, meanstructure = TRUE, estimator = "MLR")
ds2 <- semTools::discriminantValidity(object = fit,merge = TRUE)
print(ds2)

```

### CFA - CB

```{r}
#| echo: true
#| id: exp2-cfa-cb
set.seed(3110)

CB <- "
G =~ G5 + G2 + G9
L =~ L5 + L6 + L7 
"

items <- c("G5","G2","G9","L5","L6","L7")

tmp <- CFA_items(df_f,vignette = c(6,7), items=items,vignette_name="CB",bonusquestion="HAAS")
fit <- lavaan::cfa(model = CB, data = tmp, estimator = "MLR",group = "VigNro") # includes vignettes explicitly
fit.meas3 <- lavaan::fitMeasures(fit,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit.meas3)

fit_metric <- lavaan::cfa(model = CB, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = "loadings") # includes 
lavaan::lavTestLRT(fit, fit_metric)
fit_scalar <- lavaan::cfa(model = CB, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = c("loadings","intercepts")) # includes 
lavaan::lavTestLRT(fit, fit_scalar)

fit <- lavaan::cfa(model = CB, data = tmp, meanstructure = TRUE,estimator = "MLR")
ds3 <- semTools::discriminantValidity(object = fit,merge = TRUE)
print(ds3)

```

### CFA - PEP

```{r}
#| echo: true
#| id: exp2-cfa-pep
set.seed(3110)

PEP <- "
C =~ C2 + C3 + C6
X =~ X4 + X8 + X9
"

items <- c("C2","C3","C6","X4","X8","X9")

tmp <- CFA_items(df_f,vignette = c(8,9), items=items,vignette_name="PEP",bonusquestion="HAAS")
fit <- lavaan::cfa(model = PEP, data = tmp, estimator = "MLR",group = "VigNro") # includes vignettes explicitly
fit.meas4 <- lavaan::fitMeasures(fit,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit.meas4)

fit_metric <- lavaan::cfa(model = PEP, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = "loadings") # includes 
lavaan::lavTestLRT(fit, fit_metric)
fit_scalar <- lavaan::cfa(model = PEP, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = c("loadings","intercepts")) # includes 
lavaan::lavTestLRT(fit, fit_scalar)


fit <- lavaan::cfa(model = PEP, data = tmp, meanstructure = TRUE,estimator = "MLR")
ds4 <- semTools::discriminantValidity(object = fit,merge = TRUE)
print(ds4)

```

### CFA - AIA

```{r}
#| echo: true
#| id: exp2-cfa-aia
set.seed(3110)

AIA <- "
S =~ S3 + S4 + S5
I =~ I1 + I2
B =~ B1 + B3 + B5
"

items <- c("S3","S4","S5","I1","I2","B1","B3","B5")
tmp <- CFA_items(df_f,vignette = c(10,11,12), items=items,vignette_name="AIA",bonusquestion="HAAS")
fit <- lavaan::cfa(model = AIA, data = tmp, estimator = "MLR",group = "VigNro") # includes vignettes explicitly
fit.meas5 <- lavaan::fitMeasures(fit,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit.meas5)

fit_metric <- lavaan::cfa(model = AIA, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = "loadings") # includes 

# Compare configural and metric models
lavaan::lavTestLRT(fit, fit_metric)
fit_scalar <- lavaan::cfa(model = AIA, data = tmp, estimator = "MLR",group = "VigNro",
                          group.equal = c("loadings","intercepts")) # includes 
lavaan::lavTestLRT(fit, fit_scalar)


fit <- lavaan::cfa(model = AIA, data = tmp, meanstructure = TRUE,estimator = "MLR")
ds5 <- semTools::discriminantValidity(object = fit,merge = TRUE)
print(ds5)

```
## Tables for manuscript

```{r}
#| echo: true
#| id: exp2-tables
#| output: asis

cfa_table<-rbind(fit.meas1,fit.meas2,fit.meas3,fit.meas4,fit.meas5)
rownames(cfa_table)<-c("EDR","FM","CB","PEP","AIA")
print(knitr::kable(cfa_table,digits=3, caption="CFA measures of fit for constructs."))

DS1<-as.data.frame(ds1)
DS2<-as.data.frame(ds2)
DS3<-as.data.frame(ds3)
DS4<-as.data.frame(ds4)
DS5<-as.data.frame(ds5)
DS1$Episode<-"EDR"
DS2$Episode<-"FM"
DS3$Episode<-"CB"
DS4$Episode<-"PEP"
DS5$Episode<-"AIA"

ds_table<-rbind(DS1,DS2,DS3,DS4,DS5)
ds_table <- ds_table %>% relocate(Episode)

print(knitr::kable(ds_table,digits=3, caption="Discriminant validity indices for sub-constructs."))

```

## Descriptives: means and SDs and loadings

Show summary of values (means, SDs, and scores) across the sub-constructs when run jointly (CFA model with  12 sub-constructs).

```{r}
#| echo: false
#| eval: true
#| id: exp2-cfa-1st-order
set.seed(3110)

first_order <- "
# sub-factors: EDR
E =~ E1 + E6 + E7
D =~ D1 + D6 + D7
R =~ R1 + R3 + R7

# sub-factors: FM
M =~ M1 + M3 + M9
F =~ F3 + F6 + F10

# sub-factors: CB
G =~ G5 + G2 + G9
L =~ L5 + L6 + L7

# sub-factors: PEP
C =~ C2 + C3 + C6
X =~ X4 + X8 + X9

# sub-factors: AIA
S =~ S3 + S4 + S5
I =~ I1 + I2
B =~ B1 + B3 + B5
"

items <- c("E1","E6","E7","D1","D6","D7","R1","R3","R7","M1","M3","M9","F3","F6","F10","G5","G2","G9","L5","L6","L7","C2","C3","C6","X4","X8","X9","S3","S4","S5","I1","I2","B1","B3","B5")
tmp <- CFA_items(df_f,vignette = 1:12, items=items,vignette_name="ALL",bonusquestion="HAAS")
fit1 <- lavaan::cfa(model = first_order, data = tmp, meanstructure = TRUE,estimator = "MLR")
fit1.meas <- lavaan::fitMeasures(fit1,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
#print(fit1.meas)
#AVE(fit1) 
#compRelSEM(fit1,tau.eq=F)

```

For figures in the manuscript.

```{r}
#| echo: false
#| id: exp2-summarize-loadings
#| output: asis

params <- parameterEstimates(fit1, standardized = TRUE)
loadings <- params[params$op == "=~", ]

subconstructs <- params[params$op == "~~", ]
library(stringr)
key_subconstructs<-subconstructs[str_detect(subconstructs$lhs,"^[A-Z]$"),]
#key_subconstructs

edr_subs<-c("E","D","R")
edr_res <- dplyr::filter(key_subconstructs, lhs %in% edr_subs & rhs %in% edr_subs & lhs != rhs)
print(knitr::kable(edr_res,digits = 2))

fm_subs<-c("F","M")
fm_res <- dplyr::filter(key_subconstructs, lhs %in% fm_subs & rhs %in% fm_subs & lhs != rhs)
print(knitr::kable(fm_res,digits = 2))

cb_subs<-c("G","L")
cb_res <- dplyr::filter(key_subconstructs, lhs %in% cb_subs & rhs %in% cb_subs & lhs != rhs)
print(knitr::kable(cb_res,digits = 2))

pep_subs<-c("C","X")
pep_res <- dplyr::filter(key_subconstructs, lhs %in% pep_subs & rhs %in% pep_subs & lhs != rhs)
print(knitr::kable(pep_res,digits = 2))

aia_subs<-c("S","I","B")
aia_res <- dplyr::filter(key_subconstructs, lhs %in% aia_subs & rhs %in% aia_subs & lhs != rhs)
print(knitr::kable(aia_res,digits = 2))

#loadings[, c("lhs", "rhs", "est", "se", "pvalue", "std.all")]

# Create a formatted table
loading_table <- data.frame(
  Factor = loadings$lhs,
  Item = loadings$rhs,
  Estimate = round(loadings$est, 3),
  SE = round(loadings$se, 3),
  Standardized = round(loadings$std.all, 3),
  p = round(loadings$pvalue, 3)
)

#print(loading_table)

vars <- unique(loadings$rhs)

descriptives <- data.frame(
  Variable = vars,
  Mean = round(sapply(tmp[, vars], mean, na.rm = TRUE), 2),
  SD = round(sapply(tmp[, vars], sd, na.rm = TRUE), 2)
)
#print(descriptives)

final_table <- merge(
  loading_table[, c("Factor", "Item", "Standardized")],
  descriptives,
  by.x = "Item",
  by.y = "Variable",
  all.x = TRUE
)

final_table$Factor <- factor(final_table$Factor,levels = c("E","D","R","M","F","G","L","C","X","S","I","B"))

final_table <- final_table[order(final_table$Factor), ]
rownames(final_table) <- NULL
knitr::kable(final_table, digits = 2, caption = "CFA Standardized Loadings with Descriptives.")

```

## Save factor scores for emotion comparison

```{r}
#| echo: true
#| eval: true
#| id: exp2-factor-scores

# Extract factor scores from your CFA
factor_scores <- lavPredict(fit1, method = "regression")
fs <- as.data.frame(factor_scores)
fs$VigNro <- tmp$VigNro
fs$ProlificID <- tmp$ProlificID
dim(tmp)
# Save to CSV
#write.csv(fs, file = here("data","factor_scores_exp2.csv"), row.names = TRUE)

```
