---
title: "Exp. 2 - Describe Data"
toc: false
knitr:
  opts_chunk:
    fig.path: "figures/"
---


```{r}
#| echo: false
#| warning: false
#| messages: false
#| eval: true
#| id: exp2-setup

library(tidyr)
library(dplyr)
library(stringr)
library(here)
library(psych)
library(ggplot2)

```

This notebook describes the **Experiment 2 data**: preprocessing, outlier removal, demographics, mean ratings, and distributions. The data in factor study experiment 2 consists of 12 vignettes, with optimised (CFA) 35 items, that utilises between-subject design for vignettes, N=1200 (N=100 for each vignette), with a subset (≈ 50/50%) of participants answering additional questions (GEMIAC N=579, HAAS N=639).

## Load and preprocess data

```{r}
#| echo: true
#| id: exp2-load-data
#| warning: false
#| messages: false
#| eval: true
source('scr/read_EFA_qualtrics.R') # custom function
source('scr/convert_items_long.R') # custom function
source('scr/participant_reliability.R') # custom function

#fn <- "data/Episodes_September 9, 2025_14.01.csv" # N=600 GEMIAC
#d <- read_EFA_qualtrics(filename = fn,bonusquestion = "GEMIAC")
fn <- "data/Episodes_September 11, 2025_09.14.csv" # N=601-1100 HAAS
d <- read_EFA_qualtrics(filename = fn,bonusquestion = "HAAS")
df <- convert_items_long(d)
reliability <- participant_reliability(df,THRESHOLD = 0.001,verbose=TRUE)

df_f <- df %>% filter(!ProlificID %in% reliability$removed_participants)
```

### Summarise participant count across sub-experiments (vignettes)

```{r}
#| echo: true
#| id: exp2-summarise-counts
#| eval: true
#| results: asis
S <- summarise(group_by(df_f,VigNro),n=n()/35) # divide by no. of items
print(knitr::kable(S))
```


```{r}
#| echo: false
#| id: exp2-load-data-rds
#| eval: false
# If needed, load data
df_f <- readRDS("data/items35_N1200.rds")
```

```{r}
#| echo: false
#| output: asis
#| id: exp2-vignette-labels

df_f$itemCategory<-NA
df_f$itemCategory[str_detect(df_f$item_label,"E|D|R")]<-"EDR"
df_f$itemCategory[str_detect(df_f$item_label,"M|F")]<-"FM"
df_f$itemCategory[str_detect(df_f$item_label,"C|S")]<-"CB"
df_f$itemCategory[str_detect(df_f$item_label,"U|P")]<-"PEP"
df_f$itemCategory[str_detect(df_f$item_label,"I|X|B")]<-"AIA"

```

## Demographics

```{r}
#| id: exp2-demographics
# get only one row per participant
demographics <- df_f %>%
  select(ProlificID, Age, Gender, OMSI,VigNro) %>%
  distinct()
#### report
print(paste("Age (M):",round(mean(demographics$Age, na.rm=TRUE),2)))
print(paste("Age (SD):",round(sd(demographics$Age, na.rm=TRUE),2)))
print(table(demographics$Gender))/nrow(demographics)
print(table(demographics$OMSI))
demographics$OMSI_cat <- factor(demographics$OMSI,
                                levels = c("Nonmusician",
                                           "Music-loving Nonmusician",
                                           "Amateur musician",
                                           "Serious amateur musician",
                                           "Semiprofessional musician",
                                           "Professional musician"),
                                labels = c("Nonmusician",
                                           "Nonmusician",
                                           "Musician",
                                           "Musician",
                                           "Musician",
                                           "Musician"))
table(demographics$OMSI_cat)/nrow(demographics)
# subsets
table(demographics$Gender, demographics$VigNro)
chisq.test(table(demographics$Gender, demographics$VigNro))


T<-table(demographics$OMSI_cat, demographics$VigNro)
print(T)

chisq.test(table(demographics$OMSI_cat, demographics$VigNro))

summary(aov(Age ~ VigNro, data = demographics))

```


## Visualise the mean item ratings

```{r}
#| echo: true
#| id: exp2-visualise
#| fig.width: 10
#| fig.height: 10
#| fig.res: 300
source('scr/visualise_means.R')
V <- NULL
V[1]<-'EDR: Enjoyment'
V[2]<-'EDR: Distraction'
V[3]<-'EDR: Relaxation'
V[4]<-'FM: Motivation'
V[5]<-'FM: Focus'
V[6]<-'CB: Group cohesion'
V[7]<-'CB: Loneliness (reduction)'
V[8]<-'PEP: Coping'
V[9]<-'PEP: Expressing feelings'
V[10]<-'AIA: Being Moved'
V[11]<-'AIA: Curiosity'
V[12]<-'AIA: Aesthetics'

# F1 should be in Motivation, temporarily rename just for figure
df_f$item_label[df_f$item_label=="F1"] <- "M_F1"

source('scr/rename_specific_items.R') # We decided to relabel some constructs for clarity
df_fR <- rename_specific_items(df_f)

# Make into a factor
df_fR$VigNro <- factor(df_fR$VigNro,labels=V)
df_fR$Construct <- df_fR$itemCategory

```

### EDR items

```{r}
#| echo: false
#| id: exp2-vis1
#| fig.width: 10
#| fig.height: 10
#| fig.res: 300
df_fR_tmp <- dplyr::filter(df_fR,Construct=="EDR")
fig1 <- visualise_means(df_fR_tmp,Construct="EDR")
print(fig1)
```

### FM items

```{r}
#| echo: false
#| id: exp2-vis2
#| fig.width: 10
#| fig.height: 10
#| fig.res: 300
df_fR_tmp <- dplyr::filter(df_fR,Construct=="FM")
fig2 <- visualise_means(df_fR_tmp,Construct="FM")
print(fig2)
```


### CB items

```{r}
#| echo: false
#| id: exp2-vis3
#| fig.width: 10
#| fig.height: 10
#| fig.res: 300

df_fR_tmp <- dplyr::filter(df_fR,Construct=="CB")
fig3 <- visualise_means(df_fR_tmp,Construct="CB")
print(fig3)
```

### PEP items

```{r}
#| echo: false
#| id: exp2-vis4
#| fig.width: 10
#| fig.height: 10
#| fig.res: 300

df_fR_tmp <- dplyr::filter(df_fR,Construct=="PEP")
fig4 <- visualise_means(df_fR_tmp,Construct="PEP")
print(fig4)
```

### AIA items

```{r}
#| echo: false
#| id: exp2-vis5
#| fig.width: 10
#| fig.height: 10
#| fig.res: 300

df_fR_tmp <- dplyr::filter(df_fR,Construct=="AIA")
fig5 <- visualise_means(df_fR_tmp,Construct="AIA")
print(fig5)
```

### Distributions

Show rating distributions across the vignettes.

```{r}
#| echo: false
#| id: exp2-visualise2
#| fig.width: 10
#| fig.height: 10
#| fig.res: 300

g <- ggplot(df_fR, aes(y=value,fill=VigNro))+
  geom_histogram(bins=5, color="black",show.legend = FALSE)+
  facet_wrap(.~VigNro)+
  coord_flip()+
  xlab("Count")+
  ylab("Rating")+
  theme_classic()
print(g)

```
