---
title: "Exp. 2 - Higher Order Structures"
toc: false
knitr:
  opts_chunk:
    fig.path: "figures/"
---

```{r}
#| echo: false
#| warning: false
#| messages: false
#| eval: true
#| id: exp2-setup
library(tidyr)
library(dplyr)
library(stringr)
library(here)
library(psych)
library(ggplot2)
library(lavaan)
library(semTools)
```

## Load data

```{r}
#| echo: true
#| id: exp2-load-data-rds
df_f <- readRDS("data/items35_N1200.rds")
source('scr/rename_specific_items.R') # We decided to relabel some constructs for clarity
df_f <- rename_specific_items(df_f)
# F1 should be in Motivation, relabelled as M9
df_f$item_label[df_f$item_label=="F1"] <- "M9"
```

```{r}
#| echo: false
#| output: asis
#| id: vignette-labels

df_f$itemCategory<-NA
df_f$itemCategory[str_detect(df_f$item_label,"E|D|R")]<-"EDR"
df_f$itemCategory[str_detect(df_f$item_label,"M|F")]<-"FM"
df_f$itemCategory[str_detect(df_f$item_label,"G|L")]<-"CB"     # Labels updated
df_f$itemCategory[str_detect(df_f$item_label,"C|X")]<-"PEP"    # Labels updated
df_f$itemCategory[str_detect(df_f$item_label,"B|I|A")]<-"AIA"  # Labels updated

```

```{r}
#| echo: true
#| id: load-scripts
source('scr/CFA_items.R')
```

## Higher order structures

These analysyses correspond to section `Reliability, discriminant validity and convergent validity of the scales` in the manuscript.

### First-Order Model (sub-constructs)

Let's explore how the episodes operate via 12 sub-constructs.

```{r}
#| echo: true
#| eval: true
#| id: exp2-cfa-1st-order
set.seed(3110)

first_order <- "
# sub-factors: EDR
E =~ E1 + E6 + E7
D =~ D1 + D6 + D7
R =~ R1 + R3 + R7

# sub-factors: FM
M =~ M1 + M3 + M9
F =~ F3 + F6 + F10

# sub-factors: CB
G =~ G5 + G2 + G9
L =~ L5 + L6 + L7

# sub-factors: PEP
C =~ C2 + C3 + C6
X =~ X4 + X8 + X9

# sub-factors: AIA
B =~ B3 + B4 + B5
I =~ I1 + I2
A =~ A1 + A3 + A5
"

items <- c("E1","E6","E7","D1","D6","D7","R1","R3","R7","M1","M3","M9","F3","F6","F10","G5","G2","G9","L5","L6","L7","C2","C3","C6","X4","X8","X9","B3","B4","B5","I1","I2","A1","A3","A5")
tmp <- CFA_items(df_f,vignette = 1:12, items=items,vignette_name="ALL",bonusquestion="HAAS")
fit1 <- lavaan::cfa(model = first_order, data = tmp, meanstructure = TRUE,estimator = "MLR")
fit1.meas <- lavaan::fitMeasures(fit1,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit1.meas)

```


### First and Second-Order Model

Let's explore how the episodes (Constructs) operate when build as second-order model with sub-constructs.

```{r}
#| echo: true
#| id: exp2-cfa-2nd
set.seed(3110)

second_order <- "
# sub-factors: EDR
E =~ E1 + E6 + E7
D =~ D1 + D6 + D7
R =~ R1 + R3 + R7

# sub-factors: FM
M =~ M1 + M3 + M9
F =~ F3 + F6 + F10

# sub-factors: CB
G =~ G5 + G2 + G9
L =~ L5 + L6 + L7

# sub-factors: PEP
C =~ C2 + C3 + C6
X =~ X4 + X8 + X9

# sub-factors: AIA
B =~ B3 + B4 + B5
I =~ I1 + I2
A =~ A1 + A3 + A5

# 2nd order factors

EDR =~ E + D + R
FM =~ M + F
CB =~ G + L 
PEP =~ C + X
AIA =~ B + I + A  
"

items <- c("E1","E6","E7","D1","D6","D7","R1","R3","R7","M1","M3","M9","F3","F6","F10","G5","G2","G9","L5","L6","L7","C2","C3","C6","X4","X8","X9","B3","B4","B5","I1","I2","A1","A3","A5")

tmp <- CFA_items(df_f,vignette = 1:12, items=items,vignette_name="ALL",bonusquestion="HAAS")
fit2 <- lavaan::cfa(model = second_order, data = tmp, meanstructure = TRUE, estimator = "MLR")

fit2.meas <- lavaan::fitMeasures(fit2,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit2.meas)

```


### Bi-factor Model

Test whether the items load on both a general factor and their specific episode factors (sub-constructs).

```{r}
#| echo: false
#| eval: true
#| id: exp2-bifactor
set.seed(3110)

bifactor_model <- "
# general factor
g =~ E1 + E6 + E7 + D1 + D6 + D7 + R1 + R3 + R7 + M1 + M3 + M9 + F3 + F6 + F10 + G5 + G2 + G9 + L5 + L6 + L7 + C2 + C3 + C6 + X4 + X8 + X9 + B3 + B4 + B5 + I1 + I2 + A1 + A3 + A5

# sub-factors: EDR
E =~ E1 + E6 + E7
D =~ D1 + D6 + D7
R =~ R1 + R3 + R7

# sub-factors: FM
M =~ M1 + M3 + M9
F =~ F3 + F6 + F10

# sub-factors: CB
G =~ G5 + G2 + G9
L =~ L5 + L6 + L7

# sub-factors: PEP
C =~ C2 + C3 + C6
X =~ X4 + X8 + X9

# sub-factors: AIA
B =~ B3 + B4 + B5
I =~ I1 + I2
A =~ A1 + A3 + A5

# orthogonal factors
# g ~~ 0*E
# g ~~ 0*D
# g ~~ 0*R  
# g ~~ 0*M  
# g ~~ 0*F
# g ~~ 0*G
# g ~~ 0*L
# g ~~ 0*C
# g ~~ 0*X
# g ~~ 0*B
# g ~~ 0*I
# g ~~ 0*A
"

items <- c("E1","E6","E7","D1","D6","D7","R1","R3","R7","M1","M3","M9","F3","F6","F10","G5","G2","G9","L5","L6","L7","C2","C3","C6","X4","X8","X9","B3","B4","B5","I1","I2","A1","A3","A5")
tmp <- CFA_items(df_f,vignette = 1:12, items=items,vignette_name="ALL",bonusquestion="HAAS")
fit3 <- lavaan::cfa(model = bifactor_model, data = tmp, meanstructure = TRUE,orthogonal=TRUE, estimator = "MLR")

fit3.meas <- lavaan::fitMeasures(fit3,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit3.meas)

# diagnostics
#parameterEstimates(fit3)  # Any weird values
#cor(tmp, use = "complete.obs")
```

### Second-order Model only

Test whether the items work through 5 constructs only (all sub-constructs are included in each construct).


```{r}
#| echo: false
#| eval: true
#| id: exp2-secondorderonly

episodes <- "
EDR =~ E1 + E6 + E7 + D1 + D6 + D7 + R1 + R3 + R7
FM =~ M1 + M3 + M9 + F3 + F6 + F10
CB =~ G5 + G2 + G9 + L5 + L6 + L7 
PEP =~ C2 + C3 + C6 + X4 + X8 + X9
AIA =~ B3 + B4 + B5 + I1 + I2 + A1 + A3 + A5
"

items <- c("E1","E6","E7","D1","D6","D7","R1","R3","R7","M1","M3","M9","F3","F6","F10","G5","G2","G9","L5","L6","L7","C2","C3","C6","X4","X8","X9","B3","B4","B5","I1","I2","A1","A3","A5")

tmp <- CFA_items(df_f,vignette = 1:12, items=items,vignette_name="ALL",bonusquestion="HAAS")
fit4 <- lavaan::cfa(model = episodes, data = tmp, meanstructure = TRUE, estimator = "MLR")
fit4.meas <- lavaan::fitMeasures(fit4,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit4.meas)

```

### First-order, Second-order and a Generic Third-order

```{r}
#| echo: false
#| eval: true
#| id: exp2-secondorder-g

episodes2 <- "

# first-order factors

# sub-factors: EDR
E =~ E1 + E6 + E7
D =~ D1 + D6 + D7
R =~ R1 + R3 + R7

# sub-factors: FM
M =~ M1 + M3 + M9
F =~ F3 + F6 + F10

# sub-factors: CB
G =~ G5 + G2 + G9
L =~ L5 + L6 + L7

# sub-factors: PEP
C =~ C2 + C3 + C6
X =~ X4 + X8 + X9

# sub-factors: AIA
B =~ B3 + B4 + B5
I =~ I1 + I2
A =~ A1 + A3 + A5

# 2nd order factors

EDR =~ E + D + R
FM =~ M + F
CB =~ G + L 
PEP =~ C + X
AIA =~ B + I + A  

# 3rd order factor

g =~ EDR + FM + CB + PEP + AIA

"

items <- c("E1","E6","E7","D1","D6","D7","R1","R3","R7","M1","M3","M9","F3","F6","F10","G5","G2","G9","L5","L6","L7","C2","C3","C6","X4","X8","X9","B3","B4","B5","I1","I2","A1","A3","A5")

tmp <- CFA_items(df_f,vignette = 1:12, items=items,vignette_name="ALL",bonusquestion="HAAS")
fit5 <- lavaan::cfa(model = episodes2, data = tmp, meanstructure = TRUE, estimator = "MLR")
fit5.meas <- lavaan::fitMeasures(fit5,c("tli","chisq","df","pvalue", "cfi", "rmsea", "srmr"))
print(fit5.meas)

```

### Visualise Models

```{r}
#| echo: false
#| id: plot-fi
#| fig.width: 10
library(semptools)
library(semPlot)
cat("First-order\n")
p1 <- semPaths(fit1, whatLabels="est",
        sizeMan = 3.25,
        node.width = 1,
        intercepts = FALSE,
        edge.label.cex = .75,
        style = "mx",
        nCharNodes = 0, 
        nCharEdges = 0,
        mar = c(10, 5, 26, 5))
#p1b<-set_cfa_layout(p1,fcov_curve = 1.25,loading_position = .8)
#plot(p1)

cat("First and second-order\n")

p2 <- semPaths(fit2, whatLabels="est",
        sizeMan = 3.25,
        node.width = 1,
        intercepts = FALSE,
        edge.label.cex = .75,
        style = "mx",
        nCharNodes = 0, 
        nCharEdges = 0,
        mar = c(10, 5, 26, 5))
#p2b<-set_cfa_layout(p2)
#plot(p2)

cat("Bifactor\n")

p3 <- semPaths(fit3, whatLabels="est",
        sizeMan = 3.25,
        node.width = 1,
        edge.label.cex = .75,
        style = "ram",
        mar = c(10, 5, 10, 5))
#plot(p3)

cat("Second-order only\n")
p4 <- semPaths(fit4, whatLabels="est",
        sizeMan = 3.25,
        node.width = 0.5,
        intercepts = FALSE,
        edge.label.cex = 0.75,
        style = "ram",
        nCharNodes = 0, 
        nCharEdges = 0,
        mar = c(10, 5, 26, 5))
#plot(p4)

cat("1st, 2nd, and third-order\n")
p5 <- semPaths(fit5, whatLabels="est",
        sizeMan = 3.25,
        node.width = 0.5,
        intercepts = FALSE,
        edge.label.cex = 0.75,
        style = "ram",
        nCharNodes = 0, 
        nCharEdges = 0,
        mar = c(10, 5, 26, 5))
#plot(p5)
```

### Summarise Model Fits

```{r}
#| echo: false
#| id: exp2-summarise-fits
#| results: asis

fit_measures <- rbind(fit1.meas,
                      fit2.meas,
                      fit3.meas,
                      fit4.meas,
                      fit5.meas)
rownames(fit_measures) <- c("1st-order",
                             "1st and 2nd-order",
                             "Bifactor",
                             "Second-order only",
                             "1-,2-, 3rd-order")
print(knitr::kable(fit_measures,digits=2))
```

### Compare Structures

In the manuscript, we only report first-order and second-order models, as the other candidates are significantly worse (conceptually and statistically).

```{r}
#| echo: true
#| eval: true
#| id: compare-structures 

o <- compareFit(fit1, fit2)
summary(o)
```

## Construct Reliability (Composite Reliability)

Measure of internal consistency using omega from the second-order model.

```{r}
#| echo: true
#| id: exp2-reliabilty

# Calculate reliability indices:
# 0.60 questionable, 0.70 acceptable, 0.80 good, 0.90 excellent
# This is done using omega, not alpha
compRelSEM(fit1,tau.eq=F)                 # all acceptable good excellent
compRelSEM(fit2,higher = c("EDR","FM","CB","PEP","AIA"),tau.eq=F) # all acceptable good excellent
compRelSEM(fit3, tau.eq=F)               # g good but constructs are poor 
compRelSEM(fit4,tau.eq=F)                # all above 0.80
compRelSEM(fit5,tau.eq=F,higher="g")     # all above 0.80

cor_matrix <- lavInspect(fit2, "cor.lv")
print(knitr::kable(cor_matrix,digits=2,caption="Correlations between the latent factors (for sub-constructs and constructs)."))

```

## Discriminant Validity


### Average Variance Explained (traditional)

Measure of discriminant validity from the first-order model.

```{r}
#| echo: true
#| id: exp2-discriminant-validity-trad

# calculate average variance explained, 0.50+ is acceptable 
lavInspect(fit1, "cor.lv")
# First-order model
AVE(fit1)  # all above 0.50
# First and second-order
AVE(fit2)  # all above 0.50
# Bifactor
AVE(fit3,omit.factors="g")  # All below 0.50
# Second-order only
AVE(fit4)  # FM, CB and AIA above 0.50, EDR and PEP below 0.50
# 1st, 2nd, and third-order
AVE(fit5)  # all above 0.50

```

### Discriminant Validity with HTMT

Let's use the modern Heterotrait--Monotrait Ratio (HTMT) method to assess discriminant validity. It uses ratio of the correlations (Henseler, Ringlet & Sarstedt, 2015). 

#### All Sub-constructs (first-order model)


```{r}
#| echo: true
#| eval: true
#| fig.height: 6
#| fig.width: 8
#| id: other-ways1
#| output: asis

htmt_matrix <- htmt(first_order,data = tmp,htmt2 = FALSE)
print(knitr::kable(htmt_matrix,digits=3,caption='Discriminant Validity at the level of Sub-Constructs.'))

diag(htmt_matrix)=NA
round(min(htmt_matrix,na.rm=TRUE),3)
round(max(htmt_matrix,na.rm=TRUE),3)
round(median(htmt_matrix,na.rm=TRUE),3)

```

#### All Constructs (second-order model)


```{r}
#| echo: true
#| eval: true
#| fig.height: 6
#| fig.width: 8
#| id: other-ways2
#| output: asis

htmt_matrix <- htmt(episodes,data = tmp,htmt2 = FALSE)
print(knitr::kable(htmt_matrix,digits=3,caption='Discriminant Validity at the level of Constructs.'))

diag(htmt_matrix)=NA
round(min(htmt_matrix,na.rm=TRUE),3)
round(max(htmt_matrix,na.rm=TRUE),3)
round(median(htmt_matrix,na.rm=TRUE),3)

```
