---
title: "Exp. 2 - Vignettes and Emotions"
toc: false
knitr:
  opts_chunk:
    fig.path: "figures/"
---

This notebook will extract the emotion ratings for Experiment 2, where about half the participants rated HAAS adjectives and the others chosen 0 to 15 GEMIAC emotion categories (specific N reported below for both subsets). The purpose here is to offer some convergent validity measures (two common emotion measures) that can be compared to sub-constructs.


```{r}
#| echo: false
#| warning: false
#| messages: false
#| id: exp2-setup
library(tidyr)
library(dplyr)
library(stringr)
library(here)
library(ggplot2)
```

## Get HAAS data

```{r}
#| echo: false
#| warning: false
#| messages: false
#| id: exp2-rawdata

source('scr/read_EFA_qualtrics.R') # custom function
source('scr/convert_items_long.R') # custom function
source('scr/participant_reliability.R') # custom function

#fn <- "data/Episodes_September 9, 2025_14.01.csv" # N=600 GEMIAC
#d <- read_EFA_qualtrics(filename = fn,bonusquestion = "GEMIAC")
fn <- "data/Episodes_September 11, 2025_09.14.csv" # N=601-1100 HAAS
d <- read_EFA_qualtrics(filename = fn,bonusquestion = "HAAS")
df <- convert_items_long(d)
reliability <- participant_reliability(df,THRESHOLD = 0.001,verbose=FALSE) 

df_f <- df %>% filter(!ProlificID %in% reliability$removed_participants)

source('scr/rename_specific_items.R') # We decided to relabel some constructs for clarity
df_f <- rename_specific_items(df_f)
```

### HAAS

```{r}
#| echo: false
#| warning: false
#| messages: false
#| id: exp2-haas-emotions

df_f_long <- df_f %>%
   pivot_longer(
     cols = starts_with("HAAS_"),
     names_to = "Emotion",
     values_to = "Rating"
   )

haas <- c("Interested","Active","Strong","Calm","Placid","Content","Anxious","Irritable","Guilty","Bored","Sad","Tired")

haas_quadrants <- c("pos_high","pos_high","pos_high","pos_low","pos_low","pos_low","neg_high","neg_high","neg_high","neg_low","neg_low","neg_low")

df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_1$",haas[1])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_2$",haas[2])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_3",haas[3])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_4",haas[4])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_5",haas[5])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_6",haas[6])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_7",haas[7])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_8",haas[8])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_9",haas[9])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_10",haas[10])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_11",haas[11])
df_f_long$Emotion <- str_replace(df_f_long$Emotion,"HAAS_12",haas[12])

df_f_long$EmotionQ <- df_f_long$Emotion
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[1],haas_quadrants[1])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[2],haas_quadrants[2])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[3], haas_quadrants[3])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[4], haas_quadrants[4])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[5], haas_quadrants[5])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[6], haas_quadrants[6])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[7], haas_quadrants[7])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[8], haas_quadrants[8])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[9], haas_quadrants[9])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[10],haas_quadrants[10])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[11],haas_quadrants[11])
df_f_long$EmotionQ <- str_replace(df_f_long$EmotionQ,haas[12],haas_quadrants[12])

# remove items, now the values are duplicated 35 times!
df_haas <- dplyr::filter(df_f_long, item =="Gen_1")
# remove empty responses, ie. those who responded to GEMIAC
df_haas <- dplyr::filter(df_haas, !is.na(Rating) & Rating>0)

# Haas ratings were made on a 5-point scale, 0 not at all, slightly, moderately, very, extremely
# Add constructs to the data
df_haas$Construct<-"??"
df_haas$Construct[df_haas$VigNro < 4]<-"EDR"
df_haas$Construct[df_haas$VigNro == 4 | df_haas$VigNro == 5]<-"FM"
df_haas$Construct[df_haas$VigNro == 6 | df_haas$VigNro == 7]<-"CB"
df_haas$Construct[df_haas$VigNro == 8 | df_haas$VigNro == 9]<-"PEP"
df_haas$Construct[df_haas$VigNro > 9]<-"AIA"

# Labels for vignettes 1 to 12.
V <- c('Enjoyment', 'Distraction', 'Relaxation', 'EnergyControl', 'Focus', 'GroupCohesion', 'ReduceLoneliness', 'Coping', 'Express', 'Spirituality', 'Curiosity', 'Aesthetics')

# Make into factors and ensure the correct order
df_haas$VigNro<-factor(df_haas$VigNro,labels=V)
df_haas$Construct<-factor(df_haas$Construct,levels = c("EDR","FM","CB","PEP","AIA"))

```

#### Visualise

```{r}
#| echo: false
#| warning: false
#| messages: false
#| id: exp2-haas-plot

haas_summary <- df_haas %>%
  group_by(VigNro, Construct, EmotionQ) %>%
  summarise(
    Mean_Rating = mean(Rating),
    SD_Rating = sd(Rating),
    N = n(),
    SE_Rating = SD_Rating / sqrt(N/4)
  ) %>%
  ungroup()

#### Summarise valence and arousal separately
haas_summary2<- df_haas %>%
  group_by(VigNro, Construct) %>%
  summarise(
    Arousal = (sum(Rating[EmotionQ %in% c("pos_high","neg_high")]) - sum(Rating[EmotionQ %in% c("pos_low","neg_low")]))/sum(Rating[EmotionQ %in% c("pos_high","neg_high","pos_low","neg_low")]),
    Valence = (sum(Rating[EmotionQ %in% c("pos_high","pos_low")]) - sum(Rating[EmotionQ %in% c("neg_high","neg_low")]))/sum(Rating[EmotionQ %in% c("pos_high","neg_high","pos_low","neg_low")]),
    N = n(),
  ) %>%
  ungroup()

# with means and SDs
haas_summary2<- df_haas %>%
  group_by(VigNro, Construct) %>%
  summarise(
    Arousal = (mean(Rating[EmotionQ %in% c("pos_high","neg_high")]) - mean(Rating[EmotionQ %in% c("pos_low","neg_low")])),
    Valence = (mean(Rating[EmotionQ %in% c("pos_high","pos_low")]) - mean(Rating[EmotionQ %in% c("neg_high","neg_low")])),
    ArousalSD = (sd(Rating[EmotionQ %in% c("pos_high","neg_high")])),
    ValenceSD = (sd(Rating[EmotionQ %in% c("pos_high","pos_low")])),
    N = n(),
  ) %>%
  ungroup()

haas_summary2$ArousalSE <- haas_summary2$ArousalSD / sqrt(haas_summary2$N/12)
haas_summary2$ValenceSE <- haas_summary2$ValenceSD / sqrt(haas_summary2$N/12)

library(ggrepel)
g0 <- ggplot(haas_summary2, 
             aes(x = Valence, 
                 y = Arousal, 
                 color = Construct, 
                 group = Construct,
                 shape=Construct,
                 label=VigNro)) +
  geom_point(size=4) +
  geom_text_repel(show.legend = F)+
  geom_errorbar(aes(ymin = Arousal - ArousalSE, ymax = Arousal + ArousalSE), width = 0.05,alpha=0.35) +
  geom_errorbar(aes(xmin = Valence - ValenceSE, xmax = Valence + ValenceSE), width = 0.05,alpha=0.35) +
#  facet_wrap(~ Construct, scales="free_x") +
  scale_color_brewer(palette = "Set1") +
  scale_shape_manual(values = c(16,17,15,16,17))+
  geom_hline(yintercept = 0, linetype="dashed", color = "gray50")+
  geom_vline(xintercept = 1.5, linetype="dashed", color = "gray50")+ # scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, by = 1)) +
  #coord_equal()+
#  labs(title = "Mean HAAS Emotion Ratings by Vignette",
#       x = "Vignette Number",
#       y = "Mean Rating") +
  theme_classic()
 # theme(legend.position = "bottom")
print(g0)

#ggsave("figures/exp2_haas_valence_arousal.png", plot = g0, width = 6, height = 5)

g1 <- ggplot(haas_summary, aes(x = VigNro, y = Mean_Rating, color = EmotionQ, group = EmotionQ)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean_Rating - SE_Rating, ymax = Mean_Rating + SE_Rating), width = 0.2) +
  facet_wrap(~ Construct, scales="free_x") +
  scale_color_brewer(palette = "Set1") +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, by = 1)) +
  labs(title = "Mean HAAS Emotion Ratings by Vignette",
       x = "Vignette Number",
       y = "Mean Rating") +
  theme_classic() +
  theme(legend.position = "bottom")


print(g1)

#print(knitr::kable(haas_summary, digits = 2, caption = "HAAS Emotion Ratings Summary"))

haas_summary3 <- df_haas %>%
    group_by(ProlificID) %>%
  summarise(
    Arousal = (sum(Rating[EmotionQ %in% c("pos_high","neg_high")]) - sum(Rating[EmotionQ %in% c("pos_low","neg_low")]))/sum(Rating[EmotionQ %in% c("pos_high","neg_high","pos_low","neg_low")]),
    Valence = (sum(Rating[EmotionQ %in% c("pos_high","pos_low")]) - sum(Rating[EmotionQ %in% c("neg_high","neg_low")]))/sum(Rating[EmotionQ %in% c("pos_high","neg_high","pos_low","neg_low")]),
    N = n(),
  ) %>%
  ungroup()

#dim(df_haas)
#dim(haas_summary3)

```

#### Correlation of HAAS emotion ratings with the factor scores (Convergent Validity)

```{r}
#| id: exp2-factor-vs-haas
#| results: asis

# Extract factor scores from your CFA
#factor_scores <- lavPredict(fit1, method = "regression")
fs0 <- read.csv(file = here("exp2/data","factor_scores_exp2.csv"))

# HAAS: Correlate with emotion ratings
df_haas_wide <- df_haas %>%
  pivot_wider(names_from = EmotionQ, values_from = Rating)
S <- summarise(group_by(df_haas_wide,ProlificID),pos_high_M=mean(pos_high,na.rm=TRUE),pos_low_M=mean(pos_low,na.rm=TRUE),neg_high_M=mean(neg_high,na.rm=TRUE),neg_low_M=mean(neg_low,na.rm=TRUE))

print(paste("N =",length(unique(S$ProlificID))))

combined_HAAS <- merge(fs0, S, by="ProlificID")
#dim(combined_HAAS) #639
vars<-c("E","D","R","M","F","G","L","C","X","B","I","A","pos_high_M","pos_low_M","neg_high_M","neg_low_M")
matrix <- dplyr::select(combined_HAAS, all_of(vars))
cm <- cor(matrix, use="pairwise.complete.obs")
print(knitr::kable(cm, digits = 2, caption = "Correlation Matrix between Factor Scores and HAAS Emotion Ratings."))

# Just arousal and valence
combined_HAAS2 <- merge(fs0, haas_summary3, by="ProlificID")
vars<-c("E","D","R","M","F","G","L","C","X","B","I","A","Arousal","Valence")
matrix <- dplyr::select(combined_HAAS2, all_of(vars))
cm <- cor(matrix, use="pairwise.complete.obs")
print(knitr::kable(cm, digits = 2, caption = "Correlation Matrix between Factor Scores and HAAS Valence and Arousal Ratings."))

```

## Get GEMIAC data

```{r}
#| echo: false
#| id: exp2-gemiac-init
source('scr/read_EFA_qualtrics.R') # custom function
source('scr/convert_items_long.R') # custom function
source('scr/participant_reliability.R') # custom function
```


```{r}
#| echo: false
#| warning: false
#| eval: true
#| results: asis
#| id: gemiac_rawdata
fn <- "data/Episodes_September 9, 2025_14.01.csv" # N=600 GEMIAC
d <- read_EFA_qualtrics(filename = fn,bonusquestion = "GEMIAC",verbose=FALSE)
df <- convert_items_long(d)
reliability <- participant_reliability(df,THRESHOLD = 0.001,verbose=FALSE) 
df_f2 <- df %>% filter(!ProlificID %in% reliability$removed_participants)
source('scr/rename_specific_items.R') # We decided to relabel some constructs for clarity
df_f2 <- rename_specific_items(df_f2)

# No. of participants in each vignette doing GEMIAC
S <- summarise(group_by(df_f2,VigNro),n=n()/35)
print(knitr::kable(S,caption='Summary'))
# Check the N
print(sum(S$n))
length(unique(df_f2$ProlificID))
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#| messages: false
#| fig.height: 14
#| fig.width: 14
#| id: gemiac_process

#str(df_f2)
#df_f2$Emotion

library(tidyverse)

df_f3 <- df_f2 %>%
  bind_cols(
    set_names(
      map(1:16, ~ as.integer(str_detect(df_f2$Emotion, paste0("(?:^|,)", .x, "(?:,|$)")))),
      paste0("emotion_", 1:16)
    )
  )
#head(df_f3)

# fixed from Qualtrics 26/11/2025
GEMIAC<-c("Filled with wonder, amazed","Moved, touched","Enchanted, in awe","????","Inspired, enthusiastic","Energetic, lively","Joyful, wanting to dance","Powerful, strong","Full of tenderness, warmhearted", "Tense, uneasy", "Relaxed, peaceful", "Melancholic, sad", "Nostalgic, sentimental", "Indifferent, bored", "Agitated, aggressive", "None of these")

colnames(df_f3)[colnames(df_f3)=="emotion_1"]<-GEMIAC[1]
colnames(df_f3)[colnames(df_f3)=="emotion_2"]<-GEMIAC[2]
colnames(df_f3)[colnames(df_f3)=="emotion_3"]<-GEMIAC[3]
colnames(df_f3)[colnames(df_f3)=="emotion_4"]<-GEMIAC[4]
colnames(df_f3)[colnames(df_f3)=="emotion_5"]<-GEMIAC[5]
colnames(df_f3)[colnames(df_f3)=="emotion_6"]<-GEMIAC[6]
colnames(df_f3)[colnames(df_f3)=="emotion_7"]<-GEMIAC[7]
colnames(df_f3)[colnames(df_f3)=="emotion_8"]<-GEMIAC[8]
colnames(df_f3)[colnames(df_f3)=="emotion_9"]<-GEMIAC[9]
colnames(df_f3)[colnames(df_f3)=="emotion_10"]<-GEMIAC[10]
colnames(df_f3)[colnames(df_f3)=="emotion_11"]<-GEMIAC[11]
colnames(df_f3)[colnames(df_f3)=="emotion_12"]<-GEMIAC[12]
colnames(df_f3)[colnames(df_f3)=="emotion_13"]<-GEMIAC[13]
colnames(df_f3)[colnames(df_f3)=="emotion_14"]<-GEMIAC[14]
colnames(df_f3)[colnames(df_f3)=="emotion_15"]<-GEMIAC[15]
colnames(df_f3)[colnames(df_f3)=="emotion_16"]<-GEMIAC[16]

# Take only one row for each participant
df_f3_unique <- df_f3 %>%
  distinct(ProlificID, .keep_all = TRUE)

#dim(df_f3_unique)
#names(df_f3_unique)

# Add constructs (based on vignettes)
df_f3_unique$Construct<-"??"
df_f3_unique$Construct[df_f3_unique$VigNro < 4]<-"EDR"
df_f3_unique$Construct[df_f3_unique$VigNro == 4 | df_f3_unique$VigNro == 5]<-"FM"
df_f3_unique$Construct[df_f3_unique$VigNro == 6 | df_f3_unique$VigNro == 7]<-"CB"
df_f3_unique$Construct[df_f3_unique$VigNro == 8 | df_f3_unique$VigNro == 9]<-"PEP"
df_f3_unique$Construct[df_f3_unique$VigNro > 9]<-"AIA"
# Make into factor and sort the order
df_f3_unique$Construct<-factor(df_f3_unique$Construct,levels = c("EDR","FM","CB","PEP","AIA"))
#levels(df_f3_unique$Construct)

# Label vignettes
V <- c('Enjoyment', 'Distraction', 'Relaxation', 'Motivation', 'Focus', 'Group Bonding', 'Reduce Loneliness', 'Comforting', 'Expression', 'Spirituality', 'Curiosity', 'Beauty')
# Make into factor 
df_f3_unique$VigNro<-factor(df_f3_unique$VigNro,labels=V)
#levels(df_f3_unique$VigNro)

#dim(df_f3_unique)
#table(df_f3_unique$Construct)
#table(df_f3_unique$VigNro,df_f3_unique$item_label)

#### Further pivot long ------

df_f3_long <- df_f3_unique %>%
   pivot_longer(
     cols = GEMIAC,
     names_to = "Emotion_New",
     values_to = "Rating"
   )
df_gemiac<-df_f3_long

# Removes the non-selected emotions
df_gemiac <- dplyr::filter(df_f3_long, !is.na(Rating) & Rating>0)

#table(df_gemiac$Emotion_New)
#sum(table(df_gemiac$Emotion_New))

GEMIAC_summary <- df_gemiac %>%
  group_by(VigNro, Construct, Emotion_New) %>%
  summarise(
    N = n()
  ) %>%
  arrange(VigNro,Construct,desc(N)) %>%
  ungroup()


GEMIAC_summary <- df_gemiac %>%
  group_by(VigNro, Construct, Emotion_New) %>%
  summarise(
    N = n(),
    .groups = 'drop'
  ) %>%
  group_by(VigNro) %>%
  mutate(
    Total = sum(N),
    Proportion = N / Total
  ) %>%
  arrange(VigNro, Construct, desc(N)) %>%
  ungroup()

```

### GEMIAC plot

```{r}
#| echo: false
#| warning: false
#| messages: false
#| fig.height: 11
#| fig.width: 11
#| eval: true
#| results: 'asis'
#| id: gemiac_plot

#gemiac_summary$N<-gemiac_summary$N/max(gemiac_summary$N)

library(RColorBrewer)
colors_15 <- c(
  brewer.pal(6, "Set2")[1:6],
    brewer.pal(9, "Set1")
)
GEMIAC_summary$valence<-"Positive"
GEMIAC_summary$valence[GEMIAC_summary$Emotion_New %in% c("Tense, uneasy","Melancholic, sad","Indifferent, bored","Agitated, aggressive")]<-"Negative"

g2 <- ggplot(GEMIAC_summary, aes(x = VigNro, y = N, color = Emotion_New, fill=valence,shape=valence,group = Emotion_New)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ Construct, scales="free_x") +
  scale_color_manual(name="GEMIAC",values = colors_15) +
  scale_shape_manual(values = c(25,24))+
  scale_fill_manual(values = c("Positive"="white","Negative"="black"))+
#  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, by = 1)) +
  labs(title = "Mean GEMIAC Emotions by Vignette",
       x = "Vignette Number",
       y = "Count") +
  theme_classic(base_size=14) +
  theme(legend.position = "bottom")
print(g2)

# alt
g3 <- ggplot(GEMIAC_summary, 
             aes(x = Emotion_New, 
                 y = N, 
                 fill = Emotion_New, 
                 #fill = valence,
                 group = Emotion_New)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ VigNro, scales="free_x") +
  scale_fill_manual(name="GEMIAC",values = colors_15) +
  xlab("Emotion")+
  coord_flip()+
  theme_classic(base_size=14)
print(g3)

```

### Top GEMIAC Emotions for Convergent Validity Summary

Add top 3 GEMIAC emotions to the convergent validity table.

```{r}
#| results: asis
#| id: gemiac_table

gemiac_summary_top3 <- GEMIAC_summary %>%
  group_by(VigNro) %>%
  slice_max(order_by = N, n = 3) %>%
  ungroup()

print(knitr::kable(gemiac_summary_top3, digits = 2, caption = "GEMIAC Emotion Ratings Summary (Top 3 for each Vignette)."))

```
